---
title: "Data Science Careers"
subtitle: "for Behavioral and Social Science Majors"
author: "Justin Marschall | D.M. Services, Inc."
date: "April 11, 2019"
output: 
  ioslides_presentation:
    widescreen: true
    logo: images/colony_brands_logo_square.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Slides

https://github.com/justinmarschall/presentations

```{r set up, echo=FALSE}
library(tidyverse)
```

## Overview

- My career path
- Data science as fulfilling work
- How to get a job in data science
- Appendix
    - Installing R & RStudio
    - Resources to learn R
    - Publicly available data

# My Career Path

## Education

- Luther College '12
    - B.A.
    - Major: Psychology
    - Minor: Music
    - Research: Forgiveness and Health
- Iowa State University '15
    - M.S. & Certificate in Quantitative Psychology
    - Major: Psychology
    - Research: Police Interrogation Tactics and False Confessions; Quantitative Methods
    
## Job History

- D.M. Services, Inc. (2015 - Current)
    - Customer Satisfaction Data Analyst
    - Data Scientist
  
## Getting from A to B

- Major interests
    - Working with data
    - Programming
- Skills
    - Research methods
    - Statistics and modeling
    - Design of experiments
    - Diverse research background
    - Communicating to non-technical audience
    
---

<center>

<script type="text/javascript" src="https://ssl.gstatic.com/trends_nrtr/1754_RC01/embed_loader.js"></script> <script type="text/javascript"> trends.embed.renderExploreWidget("TIMESERIES", {"comparisonItem":[{"keyword":"/m/0jt3_q3","geo":"US","time":"2004-01-01 2019-04-09"}],"category":0,"property":""}, {"exploreQuery":"date=all&geo=US&q=%2Fm%2F0jt3_q3","guestPath":"https://trends.google.com:443/trends/embed/"}); </script>

</center>
    
## Common Educational Tracks for Data Science

- Statistics
- Computer Science
- Economics
- Math
- Business
- *Psychology*
- Physics
- Engineering
- "...or other related quantitative field of study"

---

<!-- http://drewconway.com/zia/2013/3/26/the-data-science-venn-diagram -->

```{r, echo=FALSE, fig.align='center', fig.height=6}
png_ds <- png::readPNG("images/Data_Science_VD.png")
grid::grid.raster(png_ds)
```

<div class="notes">

The Data Science Venn Diagram

&copy; Drew Conway Data Consulting, LLC. 2015

Used under [Creative Commons Attribution-NonCommercial License](https://creativecommons.org/licenses/by-nc/3.0/legalcode)

http://drewconway.com/zia/2013/3/26/the-data-science-venn-diagram

</div>

---

<!-- https://www.luther.edu/computer-science/data-science-major/ -->

```{r, echo=FALSE, fig.align='center', fig.height=6}
png_lc <- png::readPNG("images/LC_Data_Science.png")
grid::grid.raster(png_lc)
```

<div class="notes">

Luther College Data Science Major

https://www.luther.edu/computer-science/data-science-major/

</div>

---

<center>

<blockquote class="twitter-tweet" data-conversation="none" data-lang="en"><p lang="und" dir="ltr"> <a href="https://t.co/RHyJ0I5a4K">pic.twitter.com/RHyJ0I5a4K</a></p>&mdash; John Swain (@swainjo) <a href="https://twitter.com/swainjo/status/1108006110517174273?ref_src=twsrc%5Etfw">March 19, 2019</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

</center>

# Data Science as Fulfilling Work

## What Do You Do?

- Data wrangling
- Feature engineering
- Statistical modeling
- Report creation
- Make empirical predictions
- Design experiments
- Employee interviewing and onboarding
- Professional development
- Code development
- Auditing

## What Do You *Actually* Do?

- Help the company make good decisions (with data)

## How Do You Do That?

<!-- https://r4ds.had.co.nz/introduction.html -->

```{r, echo=FALSE, fig.align='center'}
png_r_4_ds <- png::readPNG("images/r_4_data_science.png")
grid::grid.raster(png_r_4_ds)
```

<div class="notes">

R for Data Science  

Wickham & Grolemund (2017)

https://r4ds.had.co.nz/

Used under [Creative Commons Attribution-NonCommercial-NoDerivs 3.0 License](https://creativecommons.org/licenses/by-nc-nd/3.0/us/legalcode)

</div>

## Typical Questions

- How likely is it that customer will respond to a marketing campaign?
- Given response, will customer pay their bill?
- Is customer actually who they say they are?
- What products should we market to customer?
- What are the strongest predictors of customer satisfaction?
- What factors contribute to call duration and call silence?
- What is the best way to measure phone agent quality?

## Why Data Science?

- Fast-paced work
- Always working on something new
- Access to data on massive scale
- Large impact on business
- Consult with other DS on their work
- Complex problems require high attention to detail and technical skill
- An engaging field of work that demands lifelong-learning

## Behavioral Statistics vs. Machine Learning

<div style="float: left; width: 45%;">

### **Behavioral Statistics**

- Interpretability
- Theoretical
- Common analytic tools
    - `lm()`
    - `glm()`
    - `aov()`
    - `t.test()`
</div>

<div style="float: right; width: 55%;">

### **Machine Learning**

- Predictability
- Applied
- Common analytic tools
    - `nnet::nnet()`
    - `xgboost::xgboost()`
    - `sparklyr::ml_random_forest()`
</div>

---

<!-- https://xkcd.com/1838/ -->

```{r, echo=FALSE, fig.align='center', fig.height=6}
png_xkcd <- png::readPNG("images/xkcd_machine_learning.png")
grid::grid.raster(png_xkcd)
```

<div class="notes">

XKCD Machine Learning

Used under [Creative Commons Attribution-NonCommercial 2.5 License](https://creativecommons.org/licenses/by-nc/2.5/legalcode)

https://xkcd.com/1838/

</div>

## Exciting Topics in Data Science

- Machine learning
- Deep learning
- Optimization
- Automation
- Recommendation engines
- Natural language processing

# Getting a Job

## Soft Skills

- Curiosity
- Creativity
- Competency
- Communication

## Hard Skills

- Coding/BI
    - SQL
    - R
    - Python
    - Shell
    - Excel
    - Git/GitHub
- Statistics
    - Predictive modeling
    - Machine learning
  
## How to Develop Technical Skills

- Take a class/minor/major
- Open source books (see Appendix)
- [Codecademy](https://www.codecademy.com)
- [Stack Overflow](https://stackoverflow.com/)
- [Kaggle](https://www.kaggle.com/)
- [FiveThirtyEight Riddler Puzzles](https://fivethirtyeight.com/tag/the-riddler/)
- [Twitter](https://twitter.com/)
    - \#rstats
    - \#TidyTuesday
  
## FiveThirtyEight: The Riddler

How many games would we expect to be needed to complete a best-of-seven series if each team has a 50 percent chance of winning each individual game?  

How about if one team has a 60 percent chance of winning each game?  

How about 70?

[Riddle](https://fivethirtyeight.com/features/how-much-should-you-bid-for-that-painting/)  
[Answer](https://fivethirtyeight.com/features/can-you-crack-this-squares-hidden-code/)

---

```{r}
riddler <- function(a_pr = 0.5) {
  a_wins <- 0
  b_wins <- 0
  
  while (a_wins < 4 && b_wins < 4) {
    game <- sample(c("A", "B"),
                   size = 1,
                   prob = c(a_pr, 1 - a_pr))
    
    if (game == "A") {
      a_wins <- a_wins + 1
    } else {
      b_wins <- b_wins + 1
    }
    
  }
  
  data.frame(a_pr = a_pr,
             n_games = a_wins + b_wins)
  
}
```

---

```{r}
set.seed(20190411)

riddler()
```

---

```{r, cache=TRUE}
rep(c(0.5, 0.6, 0.7), each = 2)

library(tidyverse)

set.seed(20190411)

map_dfr(rep(c(0.5, 0.6, 0.7), each = 10000),
        riddler) %>% 
  head()
```

---

```{r, cache=TRUE}
set.seed(20190411)

map_dfr(rep(c(0.5, 0.6, 0.7), each = 10000),
        riddler) %>% 
  group_by(a_pr) %>% 
  summarize(avg_games = mean(n_games),
            n = n())
```

## A Great Riddle!

- Creating user functions
- `while()` loops
- `if()` logic
- Diverse data structures
- Simulation
- Iteration
- Descriptive statistics
- Auditing code and results
- Easy to extend and visualize

---

```{r, eval=FALSE}
library(plotly)

set.seed(20190411)

# this simulation will take longer to run
map_dfr(rep(seq(0.50, 1.00, 0.01), each = 10000),
        riddler) %>% 
  group_by(a_pr) %>% 
  summarize(avg_games = mean(n_games),
            n = n()) %>% 
  plot_ly(x = ~a_pr, y = ~avg_games, type = "scatter", mode = "lines",
          text = ~avg_games, hoverinfo = "x+text") %>% 
  layout(title = "Expected Number of Games in Best-of-Seven Series 
         \nGiven Varying Win Probability",
         xaxis = list(title = "Team A Win Probability"),
         yaxis = list(title = "Games", range = c(0, 7)),
         hovermode = "compare")
```

---

<center>

```{r, echo=FALSE, fig.width=8, fig.height=6, cache=TRUE}
library(plotly)

set.seed(20190411)

# this simulation will take longer to run
map_dfr(rep(seq(0.50, 1.00, 0.01), each = 10000),
        riddler) %>% 
  group_by(a_pr) %>% 
  summarize(avg_games = mean(n_games),
            n = n()) %>% 
  plot_ly(x = ~a_pr, y = ~avg_games, type = "scatter", mode = "lines",
          text = ~avg_games, hoverinfo = "x+text") %>% 
  layout(title = "Expected Number of Games in Best-of-Seven Series \nGiven Varying Win Probability",
         xaxis = list(title = "Team A Win Probability"),
         yaxis = list(title = "Games", range = c(0, 7)),
         hovermode = "compare")
```

</center>

## Ways to Stay Up-to-Date

<div style="float: left; width: 55%;">

- Follow top packages and developers
    - [Hadley Wickham](https://github.com/hadley)
    - [Romain François](https://github.com/romainfrancois)
    - [Jenny Bryan](https://github.com/jennybc)
- Subscribe to DS blogs/podcasts
    - [R Views](https://rviews.rstudio.com/)
    - [Stack Overflow Blog](https://stackoverflow.blog/)
    - [Not So Standard Deviations](http://nssdeviations.com/)
    - [Variance Explained](http://varianceexplained.org/)
    - [Live Free or Dichotomize](https://livefreeordichotomize.com/)
    
</div>

<div style="float: right; width: 45%;">

- Attend conferences
    - [rstudio::conf](https://www.rstudio.com/conference/)
    - [useR!](https://www.r-project.org/conferences/)
    
</div>

## Advice to Get Hired

- Internship
- Graduate school
- Complete a research project outside of class
    - Faculty research
    - Web scraping
    - Publicly available data (see Appendix)
    
# Contact

## Colony Brands, Inc.

<div style="float: left; width: 38%;">
- Analytic opportunities
    - Credit risk
    - Marketing
    - Retail
    - Product
    - Fraud
    - Customer experience
- Paid summer internship
- Full-time analyst position
</div>

<div style="float: right; width: 62%;">
```{r, echo=FALSE, fig.align='left'}
png_cb <- png::readPNG("images/colony_brands_logo.png")
grid::grid.raster(png_cb, height = 0.5)
```
</div>

To apply or learn more visit [www.colonybrands.com](https://www.colonybrands.com)

## R You Ready? Taking Small Steps Toward Becoming an R Statistics User

- Talk
    - **Friday, April 12 (TOMORROW)**
    - **4:00 pm**
    - **Valders 350**
- Overview
    - Getting Started
    - Essential R: 10 Packages for the New and Intermediate UseR
    - Resources for Future Learning

## Contact
\
`r fontawesome::fa("envelope", fill = "steelblue")` justin.marschall<!-- -->@imsdm.com  
`r fontawesome::fa("twitter", fill = "steelblue")` jcmarschall  
`r fontawesome::fa("github", fill = "steelblue")` justinmarschall  
`r fontawesome::fa("linkedin", fill = "steelblue")` justinmarschall

## Slides

https://github.com/justinmarschall/presentations

# Appendix

## Installing R & RStudio

- [R](https://mirror.las.iastate.edu/CRAN/)
- [RStudio](https://www.rstudio.com/products/rstudio/download/)

## Resources to Learn R

- [R for Data Science](https://r4ds.had.co.nz/)
- [Advanced R](https://adv-r.hadley.nz/)
- [The Tidyverse Cookbook](https://rstudio-education.github.io/tidyverse-cookbook/)
- [R Graphics Cookbook](http://www.cookbook-r.com/Graphs/)
- [Learning Statistics with R](https://learningstatisticswithr.com/book/)
- [The tidyverse style guide](https://style.tidyverse.org/)
- [Applied Predictive Modeling](http://appliedpredictivemodeling.com/)

## More Resources to Learn R

- [RStudio Cheat Sheets](https://www.rstudio.com/resources/cheatsheets/)
- [RStudio Online Learning](https://www.rstudio.com/online-learning/)
- [RStudio Webinars](https://resources.rstudio.com/webinars)
- [Codecademy](https://www.codecademy.com)
- [R Vocabulary](http://adv-r.had.co.nz/Vocabulary.html)
- [swirl](https://swirlstats.com/)
- [Happy Git and GitHub for the useR](https://happygitwithr.com/)

## Publicly Available Data

- [UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/index.php)
- [Kaggle](https://www.kaggle.com/datasets)
- [Sports Reference](https://www.sports-reference.com/)
- [Quandl](https://www.quandl.com/)
- [data.gov](https://www.data.gov/)
- [FiveThirtyEight](https://data.fivethirtyeight.com/)
- Built-in data sets in R

```{r, eval=FALSE}
data()
```

## Licenses

- The Data Science Venn Diagram
    - http://drewconway.com/zia/2013/3/26/the-data-science-venn-diagram
    - [Creative Commons Attribution-NonCommercial License](https://creativecommons.org/licenses/by-nc/3.0/legalcode)
    - &copy; Drew Conway Data Consulting, LLC. 2015
- R for Data Science
    - https://r4ds.had.co.nz/
    - [Creative Commons Attribution-NonCommercial-NoDerivs 3.0 License](https://creativecommons.org/licenses/by-nc-nd/3.0/us/legalcode)
    - Wickham & Grolemund (2017)
- XKCD Machine Learning
    - https://xkcd.com/1838/
    - [Creative Commons Attribution-NonCommercial 2.5 License](https://creativecommons.org/licenses/by-nc/2.5/legalcode)


